### 2015-10-25 ch02-qlogms/logMS_IMTC.tex
Some TODOs in QLogMS

\def\thisDir{ch02-qlogms}
% \begin{abstract}
% The logarithmic distribution of spectral lines in excitation signals is widely used to measure the transfer functions of dynamic systems over a wide frequency band, covering several decades.
% Periodic signals are very popular in many advanced dynamic signal analyzers.
% Generating multitone periodic signals requires an equidistant frequency grid which conflicts with the logarithmic distribution, especially at the low frequencies.
% In this paper we offer an elegant solution to get around this problem using an improved choice for the amplitude spectrum of the multitone.
% We also offer a simple way to tune the frequency spacing of such a signal to allow for robust identification of the system under test.
% \end{abstract}

\section{Introduction} \label{sec:intro}
  It often occurs that we wish to measure the frequency response function (FRF) of a linear time invariant (LTI) system in a frequency band that covers several decades \cite{bragos2001characterisation,sanchez2011optimal,niedostatkiewicz2009,VanGheem2004}.
  Of course, we also want to attain a specified uncertainty level over the full frequency span.

  Any LTI system of finite order can be expanded into a  finite sum of systems of first and second order by means of a partial fraction expansion\cite{Oppenheim1983}.
  Therefore the performance of the proposed method is assessed on a prototype system of second order.
  This does not limit the generality of the approach as we can add several of such systems to construct a general LTI system.

@@

  The main problem is that network and dynamic signal analyzers typically
  rely on an equidistant frequency grid that comes with the Discrete Fourier Transform (DFT).
  A logarithmically spaced signal, on the other hand, does not match well to an equidistant frequency grid.
  More specifically, the lack of frequency resolution at the lowest frequencies hampers the quality of the power distribution at the low edge of the frequency band.
  To circumvent this disadvantage, we will construct quasi-logarithmically spaced signals.
  Their frequency spacing is approximately logarithmic but nevertheless conforms to the DFT's linear frequency grid.
  This problem was already discussed in \cite{GeerardynIMTC2012}.

@@

  When $\damping < \frac{1}{\sqrt{2}}$, such a second order system exhibits a peak in
  the amplitude of the transfer function \cite{Oppenheim1983}:
  \begin{align}
   \left| G\left(\wM \right) \right| & \isdef
          \max_{\omega} \left| G\left( \omega \right)  \right|
        = \frac{\dc}
               {2 \damping \sqrt{1-\damping^2}}
               \label{eq:peakAmplitude}\\
   \wM &
        = \sqrt{1-2\damping^2} \wn
        \label{eq:peakFrequency}
    \text{.}
  \end{align}

@@

\subsection{Multisine Excitations}
  To excite the system under test we will use multisine excitations \cite{Pintelon2001} with $F$ commensurate frequency components:
  \begin{equation}
    u \left( t\right) = \frac{1}{\sqrt{F}}
   \sum_{k=1}^{F} 
     A_k 
     \sin 
       \left(2\pi f_k t + \phi_k \right)
  \text{.}
  \label{eq:MultisineCT}
  \end{equation}
  In this expression, $f_k$ represents the $k^{\text{th}}$ excited frequency component.
  $A_k$ is the amplitude and $\phi_k$ the phase of the $k^{\text{th}}$ spectral line.
  We limit ourselves, without loss of generality, to random phase multisines.
  Hence, $\phi_k$ is the outcome of a uniform random process over $\left[0,2\pi\right[$.
  This leaves two sets of free parameters to create a signal that suits our needs:
    the amplitude of the spectral lines $A_k$ and
    the frequency grid $\left\{f_1, \ldots, f_F \right\}$.
  For simplicity, we will assume this grid is sorted, i.e. $f_1 < f_2 < \ldots < f_F$.

\subsection{Equidistant (Linear) Grid Multisine}
  In many applications, an equidistant frequency grid is used for $u\left( t\right) $ because this fits best with the classical DFT analysis \cite{OppenheimDT,Mandal2007}. It allows to analyze periodic signals on an equidistant frequency grid very efficiently.

@@

\subsection{Quasi-Logarithmic Grid Multisine}
  Since most measurement equipment and analysis methods rely on signals with an equidistant frequency grid, we will create a frequency grid
  $\set{f_1, f_2, f_3,\ldots, f_F}$ for which \eqref{eq:logGrid} is approximately 
  valid but with the strict constraint that the frequencies must be a subset 
  of an equidistant frequency grid. This yields the following relation for 
  subsequent frequency lines:
  \begin{equation}
    f_k \approx \alpha \cdot f_{k-1}
    \emspace \forall k \in \set{1,\ldots,F}
  \label{eq:qlogGrid}
  \end{equation}
  under the constraint that
  \begin{equation}
    f_k = N \cdot \Delta f,\; N \in \NN
    \text{.}
    \label{eq:constraintLinGrid}
  \end{equation}
  Such a grid is called a quasi-logarithmic (quasi-log) grid \cite{Pintelon2001}.

@@

  \subsection{Pink Quasi-Logarithmic Multisine}
  In \cite{Rojas2007} it is suggested that band-limited $1/f$ noise is a reasonable, albeit not optimal, robust excitation signal.

@@

  The covariance matrix $\C_{G\left(\omega_{\interest}\right)}$ of the parametric transfer function, which is pretty hard because the common lack of intuition about this model given the parameters, can be shown \cite{Pintelon2001, Gallager2008, matrixcookbook} to be approximately equal to
  \begin{align}
    \C_{G\left( \omega_{\interest}\right) } \approx
      \InnerQuadraticForm[\HT]
                         {\jac{\interest}}
                         {\FisherInformation_{\theta}^{-1}}
     \label{eq:varPeak}
  \end{align}
  with 
  \begin{equation}
    \FisherInformation_{\theta} =
    2
    \real{
          \InnerQuadraticForm[\HT]
                              {\jac{\excited}}
                              {\C_{X}}
          }
    \label{eq:FisherInfo}
  \end{equation}
  the Fisher information matrix \cite{Pintelon2001}.

@@

  Finally, $\C_{X}$ denotes the covariance matrix of $X\left(\omega\right) = G\left(\omega, \theta_0 \right) - \frac{Y_0\left(\omega \right) + E\left( \omega \right)  }{ U\left( \omega \right) } $, evaluated at the excited frequencies $\omega_{\excited}$.
  By assuming that the $X\left( \omega \right) $ are independently distributed over the frequency $\omega$, the variance-covariance matrix $\C_{X}$ is a diagonal matrix \cite{Pintelon2001}:
  \begin{equation}
    \C_{X} =
               \left[ \left.
               \begin{matrix}
                 \frac{\sigma^2_E\left(\omega_1 \right) }
                      {\abs{ U\left(\omega_1 \right)}^2 } & 0 & 0 \\
                 0 & \ddots & 0 \\
                 0 & 0 & \frac{\sigma^2_E\left(\omega_F \right) }
                              {\abs{U\left(\omega_F \right)}^2 }
               \end{matrix}
               \right.^{\vphantom{x}}\right]
    \text{.}
  \end{equation}

@@

  It is easy to show that for $m = \alpha^N$ the following relation holds:
  \begin{equation}
    G_1\left( \vec{\omega} \right)
    \stackrel{(a)}{=} G_{\alpha^N} \left( \alpha^N \vec{\omega} \right)
    \stackrel{(b)}{=} G_{\alpha^N} \left( \vec{\omega} \right)
    \label{eq:gridPeriodicity}
  \end{equation}
  since
  \begin{itemize}
   \item[(a)] the numerical values of the FRFs $G_1(\vec{\omega})$ and $G_m(m\vec{\omega})$ are equal and,
   \item[(b)] the grids $\alpha^N \vec{\omega} = \set{\alpha^{k+N} \omega_0 \given{k \in \ZN}}$ and $\vec{\omega} = \set{\alpha^k \omega_0 \given{k \in \ZN}}$ coincide for integer values of $N$.
  \end{itemize}

@@

  This form can also be obtained by applying Tustin's method or bilinear transform \cite{Oppenheim1983} to the continuous time system \eqref{eq:secondOrderSystemCT}.

@@

  The obtained FRF is then fitted to the transfer function model \eqref{eq:secondOrderSystemDTidentifiable} using the \texttt{elis} function of the \FDIDENT \cite{FDIDENT} for \matlab.
  This is a maximum likelihood estimator that fits a parametric model to the measured data in the frequency domain.
  To aid convergence and avoid local minima, the initial values for the model are chosen equal to the parameters of the true system $G_0(z)$, so that we can focus completely on the variance of the estimations.

@@

    The BK~1613 contains a set of eleven analog passive band-pass filters with center frequencies $\fc$ that are logarithmically distributed between $31.5\unit{Hz}$ and $31.5\unit{kHz}$ inclusive.
    According to its manual \cite{datasheet_bk1613}, a $0.5\unit{dB}$ ripple with three resonances is present in the pass-band and the $3\unit{dB}$ break frequencies are located at $\frac{\fc}{\sqrt{2}}$ and $\fc \sqrt{2}$.
    A first guess to realize such a filter is a third-order Chebyshev band-pass filter of Type I \cite{Zverev1967} for which the poles have a damping that is larger than $0.09$.
    Since the different pole pairs that constitute such a filter have a different damping, this is an ideally suited system to evaluate the proposed method.
    The `shape' of the transfer function of the different filters is nominally identical, which allows for an easy comparison between the different filters of the model uncertainty.

@@

    For the parametric model, on the other hand, a continuous time LTI model of the form
    \begin{equation}
      G(s) = \frac{\sum_{i=0}^3 b_i s^i}{\sum_{i=0}^6 a_i s^i}
    \end{equation}
    was estimated using \texttt{elis} \cite{FDIDENT}.
    The initial estimate was a third-order band-pass type I Chebyshev filter \cite{Zverev1967} that conformed to the information given in the datasheet \cite{datasheet_bk1613}.

>>>>

\def\thisDir{ch02-qlogms}
\TODO{reshape figures}
\TODO{rename labels}
\TODO{import specific bib items}
% \begin{abstract}
% The logarithmic distribution of spectral lines in excitation signals is widely used to measure the transfer functions of dynamic systems over a wide frequency band, covering several decades.
% Periodic signals are very popular in many advanced dynamic signal analyzers.
% Generating multitone periodic signals requires an equidistant frequency grid which conflicts with the logarithmic distribution, especially at the low frequencies.
% In this paper we offer an elegant solution to get around this problem using an improved choice for the amplitude spectrum of the multitone.
% We also offer a simple way to tune the frequency spacing of such a signal to allow for robust identification of the system under test.
% \end{abstract}

\section{Introduction} \label{sec:intro}
  It often occurs that we wish to measure the frequency response function (FRF) of a linear time invariant (LTI) system in a frequency band that covers several decades \citep{bragos2001characterisation,sanchez2011optimal,niedostatkiewicz2009,VanGheem2004}.
  Of course, we also want to attain a specified uncertainty level over the full frequency span.

  Any LTI system of finite order can be expanded into a  finite sum of systems of first and second order by means of a partial fraction expansion\citep{Oppenheim1983}.
  Therefore the performance of the proposed method is assessed on a prototype system of second order.
  This does not limit the generality of the approach as we can add several of such systems to construct a general LTI system.

@@

  The main problem is that network and dynamic signal analyzers typically
  rely on an equidistant frequency grid that comes with the Discrete Fourier Transform (DFT).
  A logarithmically spaced signal, on the other hand, does not match well to an equidistant frequency grid.
  More specifically, the lack of frequency resolution at the lowest frequencies hampers the quality of the power distribution at the low edge of the frequency band.
  To circumvent this disadvantage, we will construct quasi-logarithmically spaced signals.
  Their frequency spacing is approximately logarithmic but nevertheless conforms to the DFT's linear frequency grid.
  This problem was already discussed in \citet{GeerardynIMTC2012}.

@@

  When $\damping < \frac{1}{\sqrt{2}}$, such a second order system exhibits a peak in
  the amplitude of the transfer function \citep{Oppenheim1983}:
  \begin{align}
   \left| G\left(\wM \right) \right| & \isdef
          \max_{\omega} \left| G\left( \omega \right)  \right|
        = \frac{\dc}
               {2 \damping \sqrt{1-\damping^2}}
               \label{eq:peakAmplitude}\\
   \wM &
        = \sqrt{1-2\damping^2} \wn
        \label{eq:peakFrequency}
    \text{.}
  \end{align}

@@

  \TODO{put block diagram here}

\subsection{Multisine Excitations}
  To excite the system under test we will use multisine excitations \citep{Pintelon2001} with $F$ commensurate frequency components:
  \begin{equation}
    u \left( t\right) = \frac{1}{\sqrt{F}}
   \sum_{k=1}^{F} 
     A_k 
     \sin 
       \left(2\pi f_k t + \phi_k \right)
  \text{.}
  \label{eq:MultisineCT}
  \end{equation}
  In this expression, $f_k$ represents the $k^{\text{th}}$ excited frequency component.
  $A_k$ is the amplitude and $\phi_k$ the phase of the $k^{\text{th}}$ spectral line.
  We limit ourselves, without loss of generality, to random phase multisines.
  Hence, $\phi_k$ is the outcome of a uniform random process over $\left[0,2\pi\right[$.
  This leaves two sets of free parameters to create a signal that suits our needs:
    the amplitude of the spectral lines $A_k$ and
    the frequency grid $\left\{f_1, \ldots, f_F \right\}$.
  For simplicity, we will assume this grid is sorted, i.e. $f_1 < f_2 < \ldots < f_F$.

  \TODO{put MATLAB code for each type of signal}

\subsection{Equidistant (Linear) Grid Multisine}
  In many applications, an equidistant frequency grid is used for $u\left( t\right) $ because this fits best with the classical DFT analysis \citep{OppenheimDT,Mandal2007}. It allows to analyze periodic signals on an equidistant frequency grid very efficiently.

@@

\subsection{Quasi-Logarithmic Grid Multisine}
  Since most measurement equipment and analysis methods rely on signals with an equidistant frequency grid, we will create a frequency grid
  $\set{f_1, f_2, f_3,\ldots, f_F}$ for which \eqref{eq:logGrid} is approximately 
  valid but with the strict constraint that the frequencies must be a subset 
  of an equidistant frequency grid. This yields the following relation for 
  subsequent frequency lines:
  \begin{equation}
    f_k \approx \alpha \cdot f_{k-1}
    \emspace \forall k \in \set{1,\ldots,F}
  \label{eq:qlogGrid}
  \end{equation}
  under the constraint that
  \begin{equation}
    f_k = N \cdot \Delta f,\; N \in \NN
    \text{.}
    \label{eq:constraintLinGrid}
  \end{equation}
  Such a grid is called a quasi-logarithmic (quasi-log) grid \citep{Pintelon2001}.

@@

  \subsection{Pink Quasi-Logarithmic Multisine}
  In \citep{Rojas2007} it is suggested that band-limited $1/f$ noise is a reasonable, albeit not optimal, robust excitation signal.

@@

\TODO{zin loopt niet goed}
  The covariance matrix $\C_{G\left(\omega_{\interest}\right)}$ of the parametric transfer function, which is pretty hard because the common lack of intuition about this model given the parameters, can be shown \citep{Pintelon2001, Gallager2008, matrixcookbook} to be approximately equal to
  \begin{align}
    \C_{G\left( \omega_{\interest}\right) } \approx
      \InnerQuadraticForm[\HT]
                         {\jac{\interest}}
                         {\FisherInformation_{\theta}^{-1}}
     \label{eq:varPeak}
  \end{align}
  with 
  \begin{equation}
    \FisherInformation_{\theta} =
    2
    \real{
          \InnerQuadraticForm[\HT]
                              {\jac{\excited}}
                              {\C_{X}^{-1}}
          }
    \label{eq:FisherInfo}
  \end{equation}
  the Fisher information matrix \citep{Pintelon2001}.

@@

  Finally, $\C_{X}$ denotes the covariance matrix of $X\left(\omega\right) = G\left(\omega, \theta_0 \right) - \frac{Y_0\left(\omega \right) + E\left( \omega \right)  }{ U\left( \omega \right) } $, evaluated at the excited frequencies $\omega_{\excited}$.
  By assuming that the $X\left( \omega \right) $ are independently distributed over the frequency $\omega$, the variance-covariance matrix $\C_{X}$ is a diagonal matrix \citep{Pintelon2001}:
  \begin{equation}
    \C_{X} =
               \left[ \left.
               \begin{matrix}
                 \frac{\sigma^2_E\left(\omega_1 \right) }
                      {\abs{ U\left(\omega_1 \right)}^2 } & 0 & 0 \\
                 0 & \ddots & 0 \\
                 0 & 0 & \frac{\sigma^2_E\left(\omega_F \right) }
                              {\abs{U\left(\omega_F \right)}^2 }
               \end{matrix}
               \right.^{\vphantom{x}}\right]
    \text{.}
  \end{equation}

@@

  \TODO{formulate as lemmas, show steps}
  It is easy to show that for $m = \alpha^N$ the following relation holds:
  \begin{equation}
    G_1\left( \vec{\omega} \right)
    \stackrel{(a)}{=} G_{\alpha^N} \left( \alpha^N \vec{\omega} \right)
    \stackrel{(b)}{=} G_{\alpha^N} \left( \vec{\omega} \right)
    \label{eq:gridPeriodicity}
  \end{equation}
  since
  \begin{itemize}
   \item[(a)] the numerical values of the FRFs $G_1(\vec{\omega})$ and $G_m(m\vec{\omega})$ are equal and,
   \item[(b)] the grids $\alpha^N \vec{\omega} = \set{\alpha^{k+N} \omega_0 \given{k \in \ZN}}$ and $\vec{\omega} = \set{\alpha^k \omega_0 \given{k \in \ZN}}$ coincide for integer values of $N$.
  \end{itemize}

@@

  This form can also be obtained by applying Tustin's method or bilinear transform \citep{Oppenheim1983} to the continuous time system \eqref{eq:secondOrderSystemCT}.

@@

  The obtained FRF is then fitted to the transfer function model \eqref{eq:secondOrderSystemDTidentifiable} using the \texttt{elis} function of the \FDIDENT \citep{FDIDENT} for \matlab.
  This is a maximum likelihood estimator that fits a parametric model to the measured data in the frequency domain.
  To aid convergence and avoid local minima, the initial values for the model are chosen equal to the parameters of the true system $G_0(z)$, so that we can focus completely on the variance of the estimations.

@@

    The BK~1613 contains a set of eleven analog passive band-pass filters with center frequencies $\fc$ that are logarithmically distributed between $31.5\unit{Hz}$ and $31.5\unit{kHz}$ inclusive.
    According to its manual \citep{datasheet_bk1613}, a $0.5\unit{dB}$ ripple with three resonances is present in the pass-band and the $3\unit{dB}$ break frequencies are located at $\frac{\fc}{\sqrt{2}}$ and $\fc \sqrt{2}$.
    A first guess to realize such a filter is a third-order Chebyshev band-pass filter of Type I \citep{Zverev1967} for which the poles have a damping that is larger than $0.09$.
    Since the different pole pairs that constitute such a filter have a different damping, this is an ideally suited system to evaluate the proposed method.
    The `shape' of the transfer function of the different filters is nominally identical, which allows for an easy comparison between the different filters of the model uncertainty.

@@

    For the parametric model, on the other hand, a continuous time LTI model of the form
    \begin{equation}
      G(s) = \frac{\sum_{i=0}^3 b_i s^i}{\sum_{i=0}^6 a_i s^i}
    \end{equation}
    was estimated using \texttt{elis} \citep{FDIDENT}.
    The initial estimate was a third-order band-pass type I Chebyshev filter \citep{Zverev1967} that conformed to the information given in the datasheet \citep{datasheet_bk1613}.
